### 2025-12-02 paper/reviews/rebuttal_oopsla26.md
Comments

Thank you for your comments and questions! We start by discussing some common high-level concerns, then address all comments in order.

# Reviewer A

> Q1) Can you clarify the novelty of your paper? Does it introduce any optimization from the multi-stage programming not known before?

The main novelties are (1) identifying previously unrecognized sources of inefficiency in PBT, and (2) applying staging to address them.

Prior to this work, neither DSL abstraction overhead nor sampling costs were viewed as optimization targets in PBT. As evidence, SOTA libraries like `Base_quickcheck` use imperative features (mutable arrays, stateful instead of functional RNG) to try to improve performance, yet retain design choices (monadic combinators, relatively expensive randomness) that introduce significant overhead.
[BCP: Yes!]
While our staging techniques are well established, their application to PBT has not been explored. This paper promotes staging as a tool for PBT developers by showing it can erase abstraction overhead, yielding substantial performance improvements.
[BCP: ... but was it hard?]

<!-- Additional novelty lies in Section 3.7, where we stage type-derived generators.
Unlike most staged libraries, which require users to understand metaprogramming,
our approach is fully automatic. Since type-derived generators are constructed
at compile time, they can be staged without altering user experience. This is a
rare example of staging "for free." -->
[BCP: include this! Seems strong.]
<!-- 
[JWC can we also say that this is an instance of two-level metaprogramming? It's a metaprogram that generates metaprograms!] -->

>  Q2) Can you elaborate on the end-to-end effectiveness of your proposed 
solution?

Our evaluation uses Etna, the standard platform for 
evaluating the bug-finding speed of PBT generators. It was designed 
as a testbench for comparing generator strategies against each other, and, 
to our knowledge, no more comprehensive benchmark suite exists for this 
purpose. Our results show consistent speedups across Etna's tasks, with 
generation time translating directly to faster bug-finding.

We acknowledge that characterizing the proportion of testing time spent 
on generation across real-world PBT workloads is an interesting question 
that merits further study, and doing so would be an important research 
contribution in itself; we view this as future work outside 
the scope of this paper.
[BCP: Maybe explain why it's not easy?]

As for compilation time of emitted generator code, we do not believe this is an
issue for two reasons. First, our metaprogrammed generators compile extremely
quickly: 33ms for Bool List up to 54ms for STLC (full table after the fold).
Amortized across testing multiple properties (which usually take seconds each),
this is negligible. Second, generators change infrequently compared to code
being tested, and the generator only needs to be recompiled when it changes.
Indeed, we anticipate that a production-ready instantiation of Allegro could let
programmers cache generator code to disk, to avoid compilation at testing-time
altogether.

> Q3) What would it take to implement type-derived generators in Scala? 
Why did you decide not to implement it?

We see no fundamental barrier to implementing type-derived staged generators in
Scala.  Indeed, the scalacheck-derived project already uses Scala's reflection
features to type-derive standard (unstaged) ScalaCheck generators. Adapting this
approach to emit ScAllegro would require engineering effort, but we anticipate
no obstacles beyond that.  We chose to implement our type-derived generators in
OCaml simply because none of the authors are Scala experts: we are much more
familiar with OCaml, so we chose to focus our implementation efforts there.

# Reviewer B:

> Typos and formatting issues on p4, l379, l475, l621, l653, l718, l824, l996

Thank you! We'll fix these for the camera ready.

[BCP: IMO we can elide low-level comments and responses.]

> l119: I'm curious whether OCaml 5 might benefit from using it's native
effect handlers in place of a monadic generator DSL.

Now we are also curious about this! Very interesting idea: direct-style generators
could potentially be very performant (and more idiomatic).

> l639: Testing is all well and good, but have you not also considered
trying to construct a more rigorous proof that AllegrOCaml is correct
by program transformation / calculation?

We had not considered this, but it would be interesting to investigate.

> l751: Given that MetaOCaml is compatible with OCaml 5, why are you
still using OCaml 4?

We were not aware that MetaOCaml was now compatible with OCaml 5 -- this is good
to know!  When we began the project earlier this year, the MetaOCaml homepage
instructed users to install a version compatible with OCaml 4.14.1, which is
what we did (it has since been updated to recommend 5.3.0).
[BCP: How hard would it be to upgrade?]

> l824: The text reads a bit oddly to me at this point, as you are
talking about an experiment you performed in the past and yet you
insist on using the present tense.

Thanks for the feedback, we'll workshop this prose to make it less awkward.

> l989: You have conjectured that because GHC is set up to perform aggressive optimisations without staging then it is likely to benefit less from the kind of optimisations you exploit. It would be worth investigating to what extent this is indeed the case. Given that GHC’s QuickCheck is the canonical PBT framework it seems particularly worthwhile to perform further experiments with it. I wonder to what extent it would be possible to disable some of GHC’s aggressive inlining, both in order to assess how much it is really paying off, and to compare its robustness to your staging approach.

This is a fair point. Our conjecture is based on anecdotal evidence; one of the authors — an experienced Haskell developer — manually applied some of Allegro’s optimizations to a few Haskell generators, and they only managed to make performance worse, not better. This is nowhere near a proof, but it discouraged us from exploring that path in the short term. We would be happy to mention this anecdotal experience and/or go into more detail about wanting to do this experiment as future work

> Can we can bridge the performance gap between bespoke
highly-optimised fuzzing approaches and the much more generic PBT
libraries exemplified by QuickCheck?

This is a really interesting question, and we think that our work speaks 
directly to part of it.[BCP: Blah.] Specifically, hand-coded fuzzers that construct 
inputs satisfying specific structural and semantic constraints are performing 
precisely the same task that PBT generators are designed for. Hand-coded fuzzers 
might be faster than PBT generators if users choose to optimize them aggressively, 
but ideally those optimizations should not need to be done manually. Our hope is 
that Allegro (and other advances from the PBT literature) can make PBT generators 
fast enough that developers could use them as highly-performant fuzzers, even 
though they are expressed in a higher-level language that is easier to write 
in and reason about. The question gets a bit less clear when considering fuzzers 
that use mutation and other techniques to obtain interesting inputs, rather than 
a hand-coded program, although ideas in papers like Coverage Guided Property-Based 
Testing and Parsing Randomness may provide paths forward in those domains as well.

# Reviewer C

> 1. Can you provide an example of how the recursive generator API in Section 3.6 is used?

Yes, it's used in Figure 3 --- we can include a figure down at Section 3.6 with a small example in the camera-ready version.

> 2. How were the experimental benchmarks chosen?

We used the subset of the pre-existing Etna benchmark suite. Etna was
designed exactly for this purpose: to test the bugfinding capabilities of
competing PBT generators.  There are two benchmarks from the original Etna we
did not use. One tests Red-Black trees, and the other tests a lambda calculus
typed with System FSub, a more sophisticated type system.  We felt that the RBT eval was
redundant with the BTS eval we already used: the generators are very similar.
We chose not to use the FSub eval for two reasons. First, it had not been implemented
in OCaml, and second, we felt this would be redundant with the STLC
evals. The FSub generator is even more complex with more combinators than the STLC one, but the testing code and properties are essentially the same:
we would see an exaggerated version of the speedups in the STLC eval.

> 3. What do you mean by "semantically identical", and are the generators also syntactically similar?

Yes, the AllegrOCaml and Base_quickcheck generors are as you imagine:
essentially identical syntactically, modulo
some extra cruft for staging. We plan to submit all of our code with the artifact evaluation, but for now, we
include an example from our evals of a generator and its staged counterpart below.

By "semantically identical", we mean that, given the same random seeds, the two
generators produce the same generated output.  This property is important
because it guarantees that all of our bug-finding speedups are solely
attributable to the improved performance of staged generators, and not because
the staged generators happen to luckily create bug-squashing outputs. The
sequence of outputs is the same between the staged and unstaged versions; the staged
generators just produce them faster.

This is not an obvious or simple property to maintain:
staged and unstaged programs can look syntactically similar but behave
non-equivalently, especially in the presence of effects (for example, `let x =
print "hi" in x;x` and `let x = .<print "hi">. in .<$x;$x>. --- the former
prints once, the latter twice). Section 3.4 is essentially about ensuring that
programmers need not worry about this and can write their generators in a
manner that is syntactically as similar as possible to the non-staged versions.

[BCP: Brief reminder what this is:]
```
let rec gen ~(lo: int) ~(hi: int)  =
  let%bind sz = size in
  let should_stop = lo >= hi || sz <= 1  in
  if should_stop
    then return E
  else
    weighted_union [
      (1., return E);
      (float_of_int sz , (
        let%bind k = int_inclusive ~lo ~hi in
        let%bind v = Nat.quickcheck_generator_parameterized bst_bespoke_limits in
        let%bind left = with_size ~size_c:(sz / 2) (gen ~lo:lo ~hi:(k - 1))  in
        let%bind right = with_size ~size_c:(sz / 2) (gen ~lo:(k + 1) ~hi:hi) in
        return (T (left, k, v, right))  
      ))
    ]

let staged_quickcheck_generator (lo: int code) (hi: int code) : tree code G.t =
  recursive (.< (.~lo, .~hi ) >.) 
  (fun go lohi -> 
    let%bind sz = size in
    let%bind (lo, hi) = split_pair lohi in
    let%bind should_stop = split_bool .< .~hi <= .~lo || .~sz <= 1 >. in
    if should_stop
      then return .< E >.
    else
      weighted_union [
        (.< 1. >., return .< E >.);
        ((G.C.i2f sz), (
          let%bind k = int_inclusive ~lo ~hi in
          let%bind v = Nat.staged_quickcheck_generator_sr_t (G.C.lift bst_bespoke_limits) in
          let%bind left = with_size ~size_c:(G.C.div2 sz) (recurse go .<(.~lo, .~k - 1) >.) in
          let%bind right = with_size ~size_c:(G.C.div2 sz) (recurse go .<(.~k + 1, .~hi) >.) in
          return (.< T (.~left, .~k, .~v, .~right) >.)))
      ]
  )
```

> 4. Why was BST (Repeated Insert) not used for the ScAllegro experiments in Figure 16?

We picked a representative from each category of generators: trees, lists, and
STLC terms, representing low, medium, and high numbers of binds, respectively.
We anticipate BST (Repeated Insert) would see a slightly smaller than the other
generators speedup due to its lower number of binds, and we would be happy to
include this information in a revision. Our goal in this evaluation subsection
was not to point-by-point reproduce the AllegrOCaml evaluation in ScAllegro, but
simply to demonstrate that the performance improvments of the Allegro technique
are portable across languges.

> I found the repeated shifting of focus between OCaml and Scala to be distracting. An alternative approach might have been to describe the entirety of the approach for OCaml, and then summarize the Scala-specific differences.

Thanks for this feedback. We'll try to further isolate the discussion of Scala to just the intro and the bits of the evaluation where we evaluate ScAllegro.

> On a related note, the paper might be easier to read if there was a single table / figure that summarized the types of various values and objects used throughout the paper. I found myself making this list as I read the paper.

This is a great idea: we'll do it with the extra space in the camera-ready version.

----

Generator Compilation Times

Bool List: 33.183ms
BST Repated Insert: 45.453ms
BST Type: 46.051ms
BST Single Pass: 50.147ms
STLC Type: 44.663ms
STLC Bespoke: 53.699ms